Hybrid Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-08-20 manuscript/main.tex
uniqueness not required

\begin{corollary}
	\label{cor:orthogonal_scores}
	Given that $\{\xi^{[l]}\}_{l=1}^\infty$ is a complete orthonormal system (CONS) for $R(\mathcal{T}^*)$, the corresponding PLS scores $\rho^{[l]}$ can be equivalently expressed as $\rho^{[l]} = \langle W^{[l]}, \xi^{[l]} \rangle_{\mathbb{H}} = \langle W, \xi^{[l]} \rangle_{\mathbb{H}}$. Furthermore, the set of PLS scores $\{\rho^{[l]}\}_{l=1}^\infty$ forms an orthogonal system, that is, for any $j \neq l$, $\mathbb{E}[\rho^{[j]} \rho^{[l]}] = 0$.
\end{corollary}
Let $\beta_{\mathrm{OLS}}$ denote a minimizer of  the mean squared error $ \mathbb{E}[\|Y - \langle \beta, W \rangle_{\mathbb{H}}\|_2^2]$. It
satisfies the normal equation
\begin{equation}\label{normal_equation}
	\mathbb{E}[(Y - \langle \beta_{\mathrm{OLS}}, W \rangle_{\mathbb{H}})W] = 0 \in \mathbb{H}.
\end{equation}
If the operator $\mathcal{T}^\ast \mathcal{T}$ is invertible,  the minimizer is uniquely determined as 
$\beta_{\mathrm{OLS}} = (\mathcal{T}^\ast \mathcal{T})^{-1} \mathcal{T}^\ast Y$.
\begin{proposition}\label{proposition:nu_l}
	At the $l$-th iteration, the 
	population level least square regression coefficient of the residual response $Y^{[l]}$ on the score $\rho^{[l]}$ is defined as:
	$$\nu^{[l]} = \langle \beta_{\mathrm{OLS}}, \xi^{[l]} \rangle_{\mathbb{H}}.$$
\end{proposition}
Proof of Proposition \ref{proposition:nu_l} is provided in Appendix \ref{section:proof:proposition:nu_l}.

@@

		\subsection{Proof of Proposition \ref{proposition:nu_l}}\label{section:proof:proposition:nu_l}

>>>>

\begin{corollary}
	\label{cor:orthogonal_scores}
	Given that $\{\xi^{[l]}\}_{l=1}^\infty$ is a complete orthonormal system (CONS) for $R(\mathcal{T}^*)$, the corresponding PLS scores $\rho^{[l]}$ can be equivalently expressed as $\rho^{[l]} = \langle W^{[l]}, \xi^{[l]} \rangle_{\mathbb{H}} = \langle W, \xi^{[l]} \rangle_{\mathbb{H}}$. Furthermore, the set of PLS scores $\{\rho^{[l]}\}_{l=1}^\infty$ forms an orthogonal system, that is, for any $j \neq l$, $\mathbb{E}[\rho^{[j]} \rho^{[l]}] = 0$.
\end{corollary}

\begin{proposition} \label{prop:ols_uniqueness} Let $\beta_{\mathrm{OLS}}$ denote a minimizer of the mean squared error $\mathbb{E}[\|Y - \langle \beta, W \rangle_{\mathbb{H}}\|_2^2]$. This minimizer satisfies the normal equation: \begin{equation*} \mathbb{E}[(Y - \langle \beta_{\mathrm{OLS}}, W \rangle_{\mathbb{H}})W] = \mathbf{0} \in {\mathbb{H}}. \end{equation*}  If $\mathcal{T}^\ast \mathcal{T}$ is not invertible, $\beta_{\mathrm{OLS}}$ may not be unique, but all solutions for $\beta_{\mathrm{OLS}}$ reside in $R(\mathcal{T}^*)$, and the corresponding fitted values $Y_{\mathrm{OLS}} = \langle \beta_{\mathrm{OLS}}, W \rangle_{\mathbb{H}}$ are always unique. \end{proposition}
Proof of Proposition \ref{prop:ols_uniqueness} is provided in Appendix \ref{section:proof:prop:ols_uniqueness}.
\begin{proposition}\label{proposition:nu_l}
	At the $l$-th iteration, the 
	population level least square regression coefficient of the residual response $Y^{[l]}$ on the score $\rho^{[l]}$ is defined as:
	$$\nu^{[l]} = \langle \beta_{\mathrm{OLS}}, \xi^{[l]} \rangle_{\mathbb{H}}.$$
\end{proposition}
Proof of Proposition \ref{proposition:nu_l} is provided in Appendix \ref{section:proof:proposition:nu_l}.

@@

\subsection{Proof of Proposition \ref{prop:ols_uniqueness}}\label{section:proof:prop:ols_uniqueness}
 \begin{proof} 
 	1. Location of Solutions for $\beta_{\mathrm{OLS}}$
 	The normal equation can be rewritten using the operator $\mathcal{T}$ and its adjoint $\mathcal{T}^*$. Recall that $E[YW] = \mathcal{T}^*Y$ and $E[W \langle \beta, W \rangle_{\mathbb{H}}] = \mathcal{T}^* \mathcal{T} \beta$. Thus, the normal equation becomes: $$ \mathcal{T}^* Y - \mathcal{T}^* \mathcal{T} \beta_{\mathrm{OLS}} = \mathbf{0}_{\mathbb{H}} $$ $$ \mathcal{T}^* \mathcal{T} \beta_{\mathrm{OLS}} = \mathcal{T}^* Y $$ This equation directly shows that $\mathcal{T}^* Y$ must be in the range of $\mathcal{T}^* \mathcal{T}$, i.e., $\mathcal{T}^* Y \in R(\mathcal{T}^* \mathcal{T})$. A fundamental property of operators in Hilbert spaces is that the range of $\mathcal{T}^* \mathcal{T}$ is equal to the range of $\mathcal{T}^*$, i.e., $R(\mathcal{T}^* \mathcal{T}) = R(\mathcal{T}^*)$. Therefore, any solution $\beta_{\mathrm{OLS}}$ to the normal equation implies that $\mathcal{T}^* Y$ must reside in $R(\mathcal{T}^*)$. If $\mathcal{T}^* \mathcal{T}$ is invertible, the unique solution is $\beta_{\mathrm{OLS}} = (\mathcal{T}^* \mathcal{T})^{-1} \mathcal{T}^* Y$. As shown in the proof "The OLS Solution $(\mathcal{T}^* \mathcal{T})^{-1} \mathcal{T}^* Y$ Resides in $R(\mathcal{T}^*)$," this unique solution itself resides in $R(\mathcal{T}^*)$. If $\mathcal{T}^* \mathcal{T}$ is not invertible, its null space (kernel) is non-trivial. In this case, there are infinitely many solutions to the normal equation. However, any solution $\beta_{\mathrm{OLS}}$ still satisfies $\mathcal{T}^* \mathcal{T} \beta_{\mathrm{OLS}} = \mathcal{T}^* Y$. Since the right-hand side $\mathcal{T}^* Y$ is an element of $R(\mathcal{T}^*)$, and $R(\mathcal{T}^* \mathcal{T}) = R(\mathcal{T}^*)$, any solution $\beta_{\mathrm{OLS}}$ must have its "predictive part" (its projection onto $R(\mathcal{T}^*)$) uniquely determined. More precisely, the minimum-norm solution (obtained via the Moore-Penrose pseudoinverse) always lies in $R(\mathcal{T}^*)$, and any other solution differs from it by an element in the null space of $\mathcal{T}^* \mathcal{T}$.

 	2. Uniqueness of Fitted Values $Y_{\mathrm{OLS}}$
 	 Despite the potential non-uniqueness of $\beta_{\mathrm{OLS}}$, the fitted values $Y_{\mathrm{OLS}}$ are always unique. Let $\beta_1$ and $\beta_2$ be two different solutions to the normal equation, i.e., $\mathcal{T}^* \mathcal{T} \beta_1 = \mathcal{T}^* Y$ and $\mathcal{T}^* \mathcal{T} \beta_2 = \mathcal{T}^* Y$. Subtracting these equations gives: $$ \mathcal{T}^* \mathcal{T} (\beta_1 - \beta_2) = \mathbf{0}_{\mathbb{H}} $$ This means that the difference $(\beta_1 - \beta_2)$ belongs to the null space of $\mathcal{T}^* \mathcal{T}$, i.e., $(\beta_1 - \beta_2) \in \text{Null}(\mathcal{T}^* \mathcal{T})$. A key property in Hilbert spaces is that $\text{Null}(\mathcal{T}^* \mathcal{T}) = \text{Null}(\mathcal{T})$. Therefore, if we let $\delta = \beta_1 - \beta_2$, then $\mathcal{T}\delta = \mathbf{0}_{L^2(\Omega)}$. Now, let's examine the fitted values corresponding to $\beta_1$ and $\beta_2$: $$ Y_{\mathrm{OLS}}(\beta_1) = \langle \beta_1, W \rangle_{\mathbb{H}} = \mathcal{T}\beta_1 $$ $$ Y_{\mathrm{OLS}}(\beta_2) = \langle \beta_2, W \rangle_{\mathbb{H}} = \mathcal{T}\beta_2 $$ Consider their difference: $$ Y_{\mathrm{OLS}}(\beta_1) - Y_{\mathrm{OLS}}(\beta_2) = \mathcal{T}\beta_1 - \mathcal{T}\beta_2 = \mathcal{T}(\beta_1 - \beta_2) = \mathcal{T}\delta $$ Since $\mathcal{T}\delta = \mathbf{0}_{L^2(\Omega)}$, it follows that: $$ Y_{\mathrm{OLS}}(\beta_1) - Y_{\mathrm{OLS}}(\beta_2) = \mathbf{0}_{L^2(\Omega)} $$ Thus, $Y_{\mathrm{OLS}}(\beta_1) = Y_{\mathrm{OLS}}(\beta_2)$, demonstrating that the fitted values are unique even when the regression coefficient $\beta_{\mathrm{OLS}}$ itself is not unique. \end{proof}
		\subsection{Proof of Proposition \ref{proposition:nu_l}}\label{section:proof:proposition:nu_l}
